Supplementary Material for _A Guide to Pre-processing High-Frequency Animal Tracking Data_

### 2021-06-28 supplement/06_bat_data.Rmd
styler run
add sd and add some analyses

## Filter by covariates

No natural bounds suggest themselves, so instead we proceed to filter by covariates, since point outliers are obviously visible.

@@

First we calculate the variable `SD`.

@@

  # count long absences from the each position
  dt_recurse[, timeSinceLastVisit :=
    ifelse(is.na(timeSinceLastVisit), -Inf, timeSinceLastVisit)]
  dt_recurse[, longAbsenceCounter := cumsum(timeSinceLastVisit > 60),
    by = .(coordIdx)
  ]

@@

# get time as human readable
data_residence[, ts := as.POSIXct(time, origin = "1970-01-01")]
```

@@

@@

@@

>>>>

## Filter by system-generated error attributes

No natural bounds suggest themselves, so instead we proceed to filter by system-generated attributes of error, since point outliers are obviously visible.

@@

First we calculate the variable `SD`, which for ATLAS systems is calculated as:

$$SD = \sqrt{{VARX} + {VARY} + 2 \times {COVXY}}$$

@@

  # count long absences from the each position
  dt_recurse[, timeSinceLastVisit :=
    ifelse(is.na(timeSinceLastVisit), -Inf, timeSinceLastVisit)]
  dt_recurse[, longAbsenceCounter := cumsum(timeSinceLastVisit > 5),
    by = .(coordIdx)
  ]

@@

# get time as human readable
data_residence[, ts := as.POSIXct(time, origin = "1970-01-01")]

# get hour of day to filter for nighttime
data_residence[, hour := data.table::hour(ts)]

# filter for hour between 8pm and 5am
data_residence = atl_filter_covariates(
  data = data_residence,
  filters = "hour > 20 | hour < 5"
)
```

@@

# get mean residence time per id
data_residence[, list(mean_residence = mean(resTime),
                      sd_residence = sd(resTime)), by = "TAG"]

# get mean residence time for all bats pooled
data_residence[, list(mean_residence = mean(resTime),
                      sd_residence = sd(resTime))]

@@

# average positions per bat after removing transit points
data_residence[, list(.N), by = "TAG"][, list(mean_positions_ = mean(N))]

@@

### Duration at foraging sites

We exclude the first and last patch of each day as being roosting related, and examine how much of the total foraging time (time between the remaining first and last patch) was spent at foraging sites. It follows that the remainder of the time must have been spent in transit, or otherwise not foraging.

```{r}
# make patch summary a datatable
setDT(patch_summary)

# exclude roost
patch_summary = patch_summary[x_median > 254000, ]

# get mean and sd of duration in patches
patch_summary[, list(mean_duration = mean(duration / 60), 
                     sd_duration = sd(duration / 60))]

# assign night id
patch_summary[, c("hour", "day") := list(
  data.table::hour(as.POSIXct(time_start, origin = "1970-01-01")),
  data.table::mday(as.POSIXct(time_start, origin = "1970-01-01")))]

patch_summary[, night := 1 + c(0, cumsum(diff(hour) > 12)), by = "id"]

# get total foraging time
foraging_proportion = patch_summary[, list(
  time_total_forage = (max(time_end) - min(time_start)) / 60, 
  time_forage_site = sum(duration / 60)),
  by = c("id", "night")
  ]

# get proporion of foraging that is at a foraging site
foraging_proportion[, list(
  mean_foraging_prop = mean(time_forage_site / time_total_forage),
  sd_foraging_prop = sd(time_forage_site / time_total_forage)
)]

```
